Experiences with planning for natural language generation

### 2008-06-21 trunk/Papers/experiences/conclusion.tex
small change to introduction; changes to discussion and conclusion

>>>>

In this paper, we introduced two novel planning domains arising from
problems in natural language generation: the sentence generation domain and
the GIVE navigation domain. We also reported on the results of a number of
experiments in which we applied off-the-shelf planners to these domains.
Our main results were mixed. While modern planners do a pretty good job of
controlling the complexity of search, they also suffer from practical
problems that limit their performance in unexpected places. In particular,
in both domains the grounding step performed by both planners dominates the
time it takes to perform the search itself. The NLG community's recent
interest in planning also presents a valuable opportunity for planning
researchers: provided some of the challenges we have highlighted can be
addressed, projects like GIVE offer a constructive platform for the
planning community to showcase their techniques to a wider audience---and
to improve the quality of their tools for real-world planning tasks.

%promising, worthwhile, far-reaching, productive

%In this paper, we have introduced two novel planning domains related
%to problems in natural language generation: the sentence generation
%problem and the GIVE navigation problem.  We have then reported on a
%number of experiments in which we applied off-the-shelf planners to
%these domains.  Our main results were that while modern planners do
%pretty well in controlling the complexity of search, they suffer from
%practical problems that limit their performance in unexpected places.
%In particular, in both domains the grounding step that both planners
%perform tends to take much more time than the search itself.

### 2008-06-21 trunk/Papers/experiences/discussion.tex
small change to introduction; changes to discussion and conclusion

The positive conclusion we can draw from the experiments reported above is
that modern planners are fairly good at controlling the search for many of
the moderately-sized NLG problems we looked at. This is particularly true
for FF in the sentence generation domain, which generates nontrivial
14-word sentences in about two seconds. There is still room for
improvement, however. For instance, by comparison a special-purpose
algorithm for referring expressions could generate ten of these 14-word
sentences in a few milliseconds. Although search efficiency is not the main
performance concern in these domains, it nevertheless becomes a factor in
scenarios like Experiment~2, where the world is almost completely
unconstrained and (unsurprisingly) the search becomes much slower.

@@

We believe that if planning is to become a mature technology which
real-world users will believe in, it must find a way to deal with the
grounding problem more gracefully than current systems do.  The
research focus in developing planning algorithms has quite rightfully
been on dealing with the search, and for this it is convenient --
depending on the particular algorithm, necessary -- to ground out all
predicates and actions before starting the search itself.  However, we
think that the domains presented here are not that unusual in their
ratio of plan length against universe size.  At the end of the day,
real-world users will care about the \emph{total} runtime of a
planner, and this is more than just the search.

By and large, it is a relatively pleasant experience to come as a
``customer'' to the planning community: Thanks to the competitions, it
is easy to identify and download a fast implementation, at least for
Linux.  However, in the course of our experiments we found (and
reported) bugs in both SGPLAN and FF.  And as our experiments have
shown, there is no one silver bullet: Even planners as closely related
as FF and SGPLAN can differ significantly in their performance on
different domains.

>>>>

The positive conclusion we can draw from the experiments reported above is
that modern planners are fairly good at controlling the search for many of
the moderately-sized NLG problems we looked at. This is particularly true
for FF in the sentence generation domain, which generates nontrivial
14-word sentences in about two seconds. There is still room for
improvement, however. For instance, by comparison a special-purpose
algorithm for referring expressions could generate ten of these 14-word
sentences in a few milliseconds. Although search efficiency is not the main
performance metric in these domains, it nevertheless becomes a factor in
scenarios like Experiment~2, where the world is almost completely
unconstrained and (unsurprisingly) the search becomes much slower.

@@

The recent trend in planning research has (rightfully) focused on the
development of alorithms that control search in sophisticated ways,
resulting in a host of planners that are more powerful and more successful
than their predecessors. Depending on the particular algorithm, however,
these planners often ground out sets of predicates and actions before
starting the search itself, which has a very pronounced effect on domains
such as those described above. Since our domains are not that unusual in
their structure and composition, we hope that the lessons learnt from our
experiences can help improve the performance of current systems. We believe
that such improvements are necessary if planning is to become a more mature
technology that can offer tools to a wider community of users. At the end
of the day, real-world users will care about the \emph{total} runtime of a
planner, and this is more than just the search time.
%We offer our larger problem instances as future challenges for the
%planning community.

%We believe that if planning is to become a mature technology which
%real-world users will believe in, it must find a way to deal with the
%grounding problem more gracefully than current systems do.  The
%research focus in developing planning algorithms has quite rightfully
%been on dealing with the search, and for this it is convenient --
%depending on the particular algorithm, necessary -- to ground out all
%predicates and actions before starting the search itself.  However, we
%think that the domains presented here are not that unusual in their
%ratio of plan length against universe size.  At the end of the day,
%real-world users will care about the \emph{total} runtime of a
%planner, and this is more than just the search.

By and large, our experiences with the planning community from the point of
view of a ``customer'' (one of the authors is not a planning researcher)
have been relatively pleasant. Thanks to the planning competitions, it is
easy to identify and download a fast implementation, at least for Linux.
However, in the course of our experiments we found (and reported) bugs in
both SGPLAN and FF. We also discovered that deciding between the range of
available planners is not always straightforward. As our experiments have
shown, even planners as closely related as FF and SGPLAN can differ
significantly in their performance on different domains.

%By and large, it is a relatively pleasant experience to come as a
%``customer'' to the planning community: Thanks to the competitions, it
%is easy to identify and download a fast implementation, at least for
%Linux.  However, in the course of our experiments we found (and
%reported) bugs in both SGPLAN and FF.  And as our experiments have
%shown, there is no one silver bullet: Even planners as closely related
%as FF and SGPLAN can differ significantly in their performance on
%different domains.

### 2008-06-21 trunk/Papers/experiences/introduction.tex
small change to introduction; changes to discussion and conclusion

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
varying-sized instances of our planning domains. We also compare these
results against an ad-hoc Java implementation of GraphPlan.  Our findings
are mixed. On the one hand, it turns out that modern planners handle some
of the search problems that arise in NLG quite easily (although large
problem instances remain a challenge). On the other hand, these same
planners spend tremendous amounts of time on preprocessing. For instance,
FF spends 90\% of its runtime in the sentence generation domain on
grounding out literals and actions, and ends up slower than our version of
GraphPlan that avoids grounding. For domains like ours, which are dominated
by the number of actions and the universe size, rather than the
combinatorics of the search problem, this observation suggests that the
overall runtime of a planner could be improved by grounding more
selectively. We therefore offer these domains as challenges for the
planning community.

>>>>

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for}---planners that have been successful
on traditional benchmarks, and in the International Planning Competition
(IPC)---on a range of problem instances from our planning domains.  We also
compare these results against an ad-hoc Java implementation of GraphPlan.
Our findings are mixed. On the one hand, it turns out that modern planners
handle some of the search problems that arise in NLG quite easily (although
large problem instances remain a challenge). On the other hand, these same
planners spend tremendous amounts of time on preprocessing. For instance,
FF spends 90\% of its runtime in the sentence generation domain on
grounding out literals and actions, and ends up slower than our version of
GraphPlan that avoids grounding. For domains like ours, which are dominated
by the number of actions and the universe size, rather than the
combinatorics of the search problem, this observation suggests that the
overall runtime of a planner could be improved by grounding more
selectively. We therefore offer these domains as challenges for the
planning community.
